---
title: "Article 1: ALUES datasets"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Article 1: ALUES datasets}
  %\usepackage[UTF-8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

There are two categories of datasets available in ALUES, *land units* and *crop requirements*. 

## Input Land Units
The *land units* datasets are input land units with properties meant for evaluation for crop production. There are two regions of datasets available: *Marinduque, Philippines*; and, *Lao Cai, Vietnam*. These datasets are encoded in three characteristics: *land or soil and terrain* (LT),  *water* and *temp*. Hence, for Marinduque, the following are the datasets:

- `MarinduqueLT`
- `MarinduqueWater`
- `MarinduqueTemp`

And for Lao Cai, Vietnam, the following are the datasets:

- `LaoCaiLT`
- `LaoCaiWater`
- `LaoCaiTemp`

Sample head of the datasets are as follows:
```{r}
library(ALUES)
head(MarinduqueLT)
head(LaoCaiLT)
```
The columns of the datasets correspond to the factors or parameters measured from the land units. These parameters are used to compare to the standard values required for the target crop. The score of the comparison is referred as the suitability score.

## Crop Requirements
There are 56 crops available in ALUES all with standard properties for each target crop. Each crop are encoded into four characteristics: *terrain*, *soil*, *water* and *temp*. For example, avocado the datasets are encoded as
`AVOCADOTerrain`, `AVOCADOSoil`, `AVOCADOWater` and `AVOCADOTemp`. The list of the crop datasets can be obtained as follows:
```{r}
d <- utils::data(package = "ALUES")
alues_data <- d$results[, "Item"]
crop_data <- regmatches(alues_data, gregexpr(paste0("^[A-Z]{2,}", collapse = "|"), alues_data))
crop_data <- unique(unlist(lapply(crop_data, function(x) substr(x, 1, nchar(x)-1))))
crop_data
```
Sample crop requirement datasets are given below:
```{r}
GUAVASoil
GUAVATemp
CINNAMONTerrain
CINNAMONWater
```
Each datasets are well documented, so make sure to check it for details and descriptions of the parameters used.