---
title: "Article 5: Custom input land units"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Article 5: Custom input land units}
  %\usepackage[UTF-8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
ALUES comes with two datasets for input land units. These are the land units of Marinduque, Philippines; and, Lao Cai, Vietnam. Likely users will have their own region of interest. As such, this article will illustrate how to prepare the data.

For any region of interest, users must specify the properties of the land units into three categories:

- terrain and soil
- water
- temperature

Note that both terrain and soil factors must be specified as one dataframe, not separate. Suppose there are 5 units under study, the three characteristics can be specified as discussed in the succeeding sections.

## Terrain and soil
Suppose for terrain, the factors targetted were: Flood and SlopeD; and, suppose for soil we have CFragm and SoilDpt. The dataframe can be prepared as follows:
```{r}
terrain_input <- data.frame(
  Flood = c(1, 2, 2, 2, 3),
  SlopeD = c(3, 4, 5, 1, 2),
  CFragm = c(10, 30, 50, 60, 40),
  SoilDpt = c(45, 60, 90, 70, 30)
)
```
So that, if this input is assessed for avocado, then the suitability score for these land units computed as follows:
```{r}
library(ALUES)
AVOCADOTerrain
AVOCADOSoil
avocado_suit <- suit("avocado", terrain=terrain_input)
head(avocado_suit[["terrain"]][["Suitability Score"]])
head(avocado_suit[["terrain"]][["Suitability Class"]])
head(avocado_suit[["soil"]][["Suitability Score"]])
head(avocado_suit[["soil"]][["Suitability Class"]])
```

## Water
For water characteristics, suppose the average rainfall for the first month (WmAv1) is 90, for second month (WmAv2) is 110, third month (WmAv3) is 150, and fourth month (WmAv4) is 160. This can be specified as follows:
```{r}
water_input <- data.frame(
  WmAv1 = c()
)
```