---
title: "Getting Started"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Getting Started}
  %\usepackage[UTF-8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Show me the code
To evaluate the soil characteristics of the land units of Marinduque, Philippines, for farming banana, use the `suit` function as follows:
```{r}
library(ALUES)
suit_banana <- suit("banana", terrain=MarinduqueLT)
head(suit_banana[["soil"]]$`Suitability Score`)
head(suit_banana[["soil"]]$`Suitability Class`)
```
To compute the overall suitability of the land units by averaging across factors, use the `overall_suit` as follows:
```{r}
osuit <- overall_suit(suit_banana[["soil"]], method="average")
head(osuit)
```
## Show me the speed
The elapsed time for computing the suitability scores and classes were recorded for the land units of Marinduque, which has 881 units (or observations) in total; and, for the region of Lao Cai, Vietnam, which has 2928 land units.
```{r}
library(microbenchmark)
microbenchmark(
  suppressWarnings(suit("banana", terrain=MarinduqueLT, interval="unbias"))
)
microbenchmark(
  suppressWarnings(suit("banana", terrain=LaoCaiLT, interval="unbias"))
)
```