---
title: "Getting Started"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Getting Started}
  %\usepackage[UTF-8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Datasets

## Input Land Units
In calculating land suitability, there are two main inputs: *land units characteristics* and *target crop requirements*. Users can input their own land units characteristics as data frame, by following the format of the land units data in ALUES, which are `MarinduqueLT` (for land and soil characteristics), `MarinduqueWater` (for water characteristics), `MarinduqueTemp` (for temperature characteristics). Another input land units data is also available in ALUES, from the province of Vietnam, Lao Cai (`LaoCaiLT`, `LaoCaiWater`, and `LaoCaiTemp`). 

## Crop Requirements
Another required input for computing the suitability of the land is the target crop requirements. In ALUES, there are 56 crops each encoded as 4 data frames, for the four characteristics, namely terrain, soil, water and temperature. For example, for avocado, the corresponding data frames are `AVOCADOTerrain`, `AVOCADOSoil`, `AVOCADOWater` and `AVOCADOTemp`. The suffixes (`Terrain`, `Soil`, `Water`, `Temp`) are consistent across crops, and and the name of the crops are consistently uppercase. So that, the naming convention is formulated as `CROPNAME` + `Suffix`.

## Computing Suitability 
To compute the suitability, the `suit` function is used. For example, to compute the suitability of the Marinduque land units for rainfed bunded rice is done as follows
```{r}
library(ALUES)
ricebr_suit <- suit("ricebr", terrain=MarinduqueLT)
```
Check the available names of the output
```{r}
names(ricebr_suit)
```
Extract scores for terrain, if any
```{r}
ricebr_suit[["terrain"]]
```
Extract scores for soil
```{r}
lapply(ricebr_suit[["soil"]], function(x) head(x))
```

## Computing  Overall Suitability  
Overall suitability can be computed across factors evluated. Example, for soil suitability computed above, the overall suitability is computed as follows
```{r}
head(overall_suit(ricebr_suit[["soil"]]))
```
There are three methods for aggregating the factors to obtain the overall suitability, and these are `minimum` (default), `average` and `maximum`.
```{r}
head(overall_suit(ricebr_suit[["soil"]], "average"))
head(overall_suit(ricebr_suit[["soil"]], "maximum"))
```